A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.

### 2014-02-23 jacobian_calc_CMAME.tex
continuing I.2

\note[Michael]{re: I.2, John I think you would better handle expanding this section,
I was going to essentially copy silling:psa.}\change[Michael]{Briefly, peridynamics}Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
is a nonlocal reformulation of the partial-differential equations that provide
the statement of momentum balance in classical continuum mechanics. Its primary
goal is to avoid the use of spatial derivatives in the balance equations or
constitutive laws such that discontinuous displacements, i.e. cracks, are
mathematically consistent with the governing equations. It has demonstrated
great promise in modeling problems with pervasive failure
\cite{littlewood2010}, interesting phenomenon such as dynamic crack branching
\cite{ha2010sod}, and is being extended to multiphyics phenomena
\cite{bobaru2011peridynamic,katiyar2013} and the problem of multi-scale
coupling to molecular simulations \cite{seleson2009peridynamics}. The related mathematical
definitions and parts of peridynamic theory especially relevant to the present
work are explained in the sequel.

The type of material model used in this study is referred to as a linear
peridynamic solid, having what is called a state-based constitutive model. In
order for a definition of this material to be useful her. To begin with, some essential
mathematical definitions will be listed and peridynamic theory summarized. Then
state-based constitutive models will be summarized, followed by a description of
the linear peridynamic solid model used in this work. After the peridynamic
equation of motion and consitutive model in use have been discussed, it is then
possible to give a more precise definition of the tangent stiffness matrix for
the problem being discussed in this work. The following discussion is largely
taken from \cite{silling:psa} but reproduced here for clarity.

\note[Michael]{re: I.2. John, this was to be the skeleton for the new sections on peridynamics.}
\subsubsection{Vector states for state-based peridynamic theory.}

\subsubsection{The state-based peridynamic equation of motion.}

>>>>

\note[Michael]{re: I.2, John I think you would better handle expanding this section,
I was going to essentially copy silling:psa.}\change[Michael]{Briefly, peridynamics}Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
is a nonlocal reformulation of the partial-differential equations that provide
the statement of momentum balance in classical continuum mechanics. Its primary
goal is to avoid the use of spatial derivatives in the balance equations or
constitutive laws such that discontinuous displacements, i.e. cracks, are
mathematically consistent with the governing equations. It has demonstrated
great promise in modeling problems with pervasive failure
\cite{littlewood2010}, interesting phenomenon such as dynamic crack branching
\cite{ha2010sod}, and is being extended to multiphyics phenomena
\cite{bobaru2011peridynamic,katiyar2013} and the problem of multi-scale
coupling to molecular simulations \cite{seleson2009peridynamics,seleson2014atom}.

The type of material model used in this study is referred to as a linear
peridynamic solid, having what is called a state based constitutive model. For
the convenience of the reader, vector states will be introduced. Then the state
based peridynamic equation of motion will be shown, followed by a description
of the linear peridynamic solid model used in this work. After the peridynamic
equation of motion and a consitutive model have been discussed, a definition of
the tangent stiffness matrix is given. The following discussion is largely
taken from \cite{silling:psa} but reproduced here for clarity.

\note[Michael]{re: I.2. John, this was to be the skeleton for the new sections on peridynamics.}
\subsubsection{Vector states for state based peridynamic theory.}
Because the vector state, which is fair to say an uncommon notation, is used in the literature discussing state based peridynamic theory
it will be enumerated here for the sake of continuity. 
%
Paraphrasing, let a spherical region, centered about a
location in threespace, called a horizon or neighborhood with a certain radius
be defined. A vector state of order $m$ is a mapping of a vector contained
within the horizon to a tensor of order $m$ \cite{silling:psa}. Vector states
have have their own mathematical properties and definitions for sums, composition,
point product, reference state, identity scalar state, null reference state, null scalar state,
dot product, magnitude state, unit state, and direction state which are detailed in \cite{silling:psa}. 
It is important to think of vector states as non-linear, not necessarily continuous maps of vectors to real Euclidian space, where second order tensors are linear maps of three dimensional vectors to threespace. 
%
One important tool mechanics gains from the concept of vector states is a way
to represent complicated deformations within a body in the form of a state
function. Whereas a symmetric second order tensor maps a spherical region in a
body in the reference configuration to an ellipsoid in the deformed
configuration, a vector state can map a spherical region in a body into a
complicated shape \cite{silling:psa}.

(PLACE PERIDYNAMIC POTATO DIAGRAM HERE)

%
The properties and definitions of vector states allow interaction with vectors
and second order tensors to be defined which includes a tensor vector product
in terms of a state function called an expansion, the tensor product of two states in terms of the
volume integral over a horizon of the dyadic product of those states times a
scalar function, the shape tensor as the tensor product of the reference vector
state with itself, and a tensor reduced form of a state as the conventionally
defined tensor product of the state tensor product of the state and the
reference state with the inverse of the shape tensor, and the inheritance of
the properties of a proper orthogonal second order tensor in its state
expansion. These ideas are detailed in \cite{silling:psa}. 
%
Because of its importance in this work, the definition of a derivative of a vector state
will be given a more complete explanation.

\subsubsection{The state based peridynamic equation of motion.}
